---
title: "Data Analytics II: Project Part A"
author: |
    | your group number here
    | your names here
    | your reg numbers here
date: "2023-XX-25"
output: pdf_document
df_print: paged
geometry: margin=2cm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(readxl)
library(kableExtra)
library(dplyr)
library(ggplot2)
library(tibble)
```

## A.1 - Explorative Data Analysis (EDA)


```{r A.1.1}

```

[your comments here]

```{r A.1.2}

```

[your comments here]

```{r A.1.3}

```

Do the variables Income and Age appear to be (approximately) normally distributed? 

Can you think of a suitable distribution for the variable children? 


## A.2 - Confidence Intervals

```{r A.2.1}

```

[your comments here]

```{r A.2.2}

```

[your comments here]

```{r A.2.3}

```

Are the underlying assumptions for calculating this interval met in this data? 

## A.3 - Confidence Intervals and the Sample Size

```{r A.3.1}

```

[your comments here]

```{r A.3.2}

```

How should we adjust the confidence level as the size of the sample increases? Why?


## A.4 - Confidence Intervals: Comparison of Population Means

```{r A.4.i}

```

[your comments here]

Are underlying assumptions used to calculate this interval met in this data?

