---
title: "Data Analytics II: Project Part A"
author: |
    | 57
    | Gabriela Siren, Didrik Gentili, Anna Gao, Ludwig Fredriksson
    | 25601, 25550, 25811, XXXXX
date: "2023-01-29"
output: pdf_document
df_print: paged
geometry: margin=2cm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(readxl)
library(kableExtra)
library(ggplot2)
library(tibble)
library(rmarkdown)
library(knitr)
library(tidyverse)
library(tinytex)
```

## A.1 - Explorative Data Analysis (EDA)


```{r A.1.1}
data <- read_excel("SegmentData57.xlsx")
stats <- function(x){c("Mean"=mean(x)
                       ,"Standard Deviation"=sd(x)
                       ,"Maximum"=max(x)
                       ,"Median"=median(x)
                       ,"Minimum"= min(x))}
summaryStatsTable <-sapply(c(data[1:1], data[3:4]),stats)
kable(summaryStatsTable, digits = 2, col.names= c("Age","Income", "Kids"))

#age <- data$age
#meanage <- mean(age)
#sdage <- sd(age)
#gender <- data$gender
#income <- data$income
#kids <- data$kids
#homes <- data$ownHome
#subs <- data[,6]
#segment <- data[,7]

```

[your comments here]

```{r A.1.2}

```

[your comments here]

```{r A.1.3}

```

Do the variables Income and Age appear to be (approximately) normally distributed? 

Can you think of a suitable distribution for the variable children? 


## A.2 - Confidence Intervals

```{r A.2.1}

```

[your comments here]

```{r A.2.2}

```

[your comments here]

```{r A.2.3}

```

Are the underlying assumptions for calculating this interval met in this data? 

## A.3 - Confidence Intervals and the Sample Size

```{r A.3.1}

```

[your comments here]

```{r A.3.2}

```

How should we adjust the confidence level as the size of the sample increases? Why?


## A.4 - Confidence Intervals: Comparison of Population Means

```{r A.4.i}

```

[your comments here]

Are underlying assumptions used to calculate this interval met in this data?

