---
title: "Data Analytics II: Project Part A"
author: |
    | 57
    | Gabriela Siren, Didrik Gentili, Anna Gao, Ludwig Fredriksson
    | 25601, 25550, 25811, XXXXX
date: "2023-01-29"
output: pdf_document
df_print: paged
geometry: margin=2cm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(readxl)
library(kableExtra)
library(ggplot2)
library(tibble)
library(rmarkdown)
library(knitr)
library(tidyverse)
library(tinytex)
library(dplyr)
```


```{r setup Importing data}
data <- read_excel("SegmentData57.xlsx")
```

## A.1 - Explorative Data Analysis (EDA)

```{r A.1.1 Statistics table: numerical variable and proportion}
stats <- function(x){c("Mean " = mean(x)
                       ,"Stan. Deviaton " = sd(x)
                       ,"Minimum " = min(x)
                       , "Q1 " = quantile(x,probs=0.25)
                       ," Median " = median(x)
                       , "Q3 " = quantile(x, probs = 0.75)
                       ,"Maximum " = max(x))}
summaryStatsTable <-sapply(c(data[1:1], data[3:4]),stats)
# I calculated the values and then input them since we only have to upload the pdf and not our code

# I added some code to calculate the category data, but im curious about how you calculated H0 and H1, i dont get the same answers


homedata = data[data$ownHome == "ownYes",]
subdata = data[data$subscribe == "subYes",]

ownHomem = sum(homedata$gender =="Male") /nrow(homedata)
ownHomef = sum(homedata$gender =="Female") /nrow(homedata)

subm = sum(subdata$gender =="Male") /nrow(subdata)
subf = sum(subdata$gender =="Female") /nrow(subdata)
print(subm)



H0 <- c(0.5233, 0.51667, 0.86667)
H1 <- c(0.47667, 0.48333, 0.13333)




categories <- rbind(H0,H1)
rownames(categories) = c("Female/False", "Male/True")

kables(
  list(
    kable(
      summaryStatsTable, digits = 2, col.names= c("Age","Income", "Kids"), valign = "t", booktabs = "True"
      )
  ,
    kable(
      categories, col.names = c("Gender","OwnsHome", "Subscribes"), valign = "t", booktabs = "True"
      )
  ),
  caption = "Summary table for all variables",
)
```
This table tells me...

```{r A.1.1 Numerical variables by segment} 
ageSM <- data %>%
  group_by(Segment) %>%
    summarise("Mean"=mean(age),
              "StDev"= sd(age),
               "Minimum"= min(age),
               "Maximum"= max(age),
              "Q1" = quantile(age,probs=0.25),
              "Median"= median(age),
               "Q3" = quantile(age,probs=0.75))
kable(ageSM, digits = 2, caption = "Summary table for age")






incomeSM <- data %>%
  group_by(Segment) %>%
    summarise("Mean"=mean(income),
              "StDev"= sd(income),
               "Minimum"= min(income),
               "Maximum"= max(income),
              "Q1" = quantile(income,probs=0.25),
              "Median"= median(income),
               "Q3" = quantile(income,probs=0.75))
kable(incomeSM, digits = 2, caption = "Summary table for income")

kidsSM <- data %>%
  group_by(Segment) %>%
    summarise("Mean"=mean(kids),
              "StDev"= sd(kids),
               "Minimum"= min(kids),
               "Maximum"= max(kids),
              "Q1" = quantile(kids,probs=0.25),
              "Median"= median(kids),
               "Q3" = quantile(kids,probs=0.75))
kable(kidsSM, digits = 2, caption = "Summary table for children")

```

[your comments here]

```{r A.1.2 Histograms}
d1 <- data$age
d2 <- data$income
d3 <- data$kids

hist(d1, breaks=30, main = "Histogram of Age", xlab= "Aggregate age")

hist(d2, breaks=100, main = "Histogram of Income", xlab= "Aggregate income")

hist(d3, breaks=5, main = "Histogram of Kids", xlab= "Aggregate kids")

```

```{r A.1.2 Boxplots}
#Aggregate income boxplot
boxplot(d2,
        main = "Aggregate Income boxplot",
        xlab = "Income",
        horizontal = TRUE,
        las=1)

#Segmented income boxplot
seg_sm<-data[data$Segment=="Suburb mix",]
seg_uh<-data[data$Segment=="Urban hip",]
seg_t<-data[data$Segment=="Travelers",]
seg_mu<-data[data$Segment=="Moving up",]

boxplot(seg_sm$income,
        main = "Segmented Income boxplot - Suburb Mix",
        xlab = "Income",
        horizontal = TRUE,
        las=1)
boxplot(seg_uh$income,
        main = "Segmented Income boxplot - Urban Hip",
        xlab = "Income",
        horizontal = TRUE,
        las=1)
boxplot(seg_t$income,
        main = "Segmented Income boxplot - Travelers",
        xlab = "Income",
        horizontal = TRUE,
        las=1)
boxplot(seg_mu$income,
        main = "Segmented Income boxplot - Moving Up",
        xlab = "Income",
        horizontal = TRUE,
        las=1)
```


[your comments here]

```{r A.1.3}

```

Do the variables Income and Age appear to be (approximately) normally distributed? 

Can you think of a suitable distribution for the variable children? 


## A.2 - Confidence Intervals

```{r A.2.1}
# R program to determine the mean

# Calculate the mean of the Segment.Data57

mean_value <- mean(data$income)

# R program to determine the mean

# Calculate the mean of the Segment.Data57

mean_value <- mean(data$income)


```

[your comments here]

```{r A.2.2}

```

[your comments here]

```{r A.2.3}

```

Are the underlying assumptions for calculating this interval met in this data? 

## A.3 - Confidence Intervals and the Sample Size

```{r A.3.1}

```

[your comments here]

```{r A.3.2}

```

How should we adjust the confidence level as the size of the sample increases? Why?


## A.4 - Confidence Intervals: Comparison of Population Means

```{r A.4.i}
travelersdata = data[data$Segment == "Travelers",]
Urban_hipdata = data[data$Segment == "Urban hip",]

#n1 = nrow(travelersdata)
#n2 = nrow(Urban_hipdata)

# 1 is for travelers 2 is for urban_hip


#m1 <- mean(travelersdata$income)
#s1 <- sd(travelersdata$income)
#t1 <- qt(0.950,n1-1)

#ME1 <- t1*s1/sqrt(n1)

#LCL1 <- m1-ME1
#UCL1 <- m1+ME1

t.test(travelersdata$income, conf.level = 0.90)
t.test(Urban_hipData$income, conf.level = 0.90)








```

[your comments here]

Are underlying assumptions used to calculate this interval met in this data?

