---
title: "Data Analytics II: Project Part A"
author: |
    | 57
    | Gabriela Siren, Didrik Gentili, Anna Gao, Ludwig Fredriksson
    | 25601, 25550, 25811, 25632
date: "2023-01-29"
output: pdf_document
df_print: paged
geometry: margin=2cm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(readxl)
library(kableExtra)
library(ggplot2)
library(tibble)
library(rmarkdown)
library(knitr)
library(tidyverse)
library(tinytex)
library(dplyr)
library(broom)
library(purrr)
```


```{r setup Importing data}
data <- read_excel("SegmentData57.xlsx")
```

## A.1 - Explorative Data Analysis (EDA)

```{r A.1.1 Statistics table: numerical variable and proportion}
stats <- function(x){c("Mean " = mean(x)
                       ,"Stan. Deviaton " = sd(x)
                       ,"Minimum " = min(x)
                       , "Q1 " = quantile(x,probs=0.25)
                       ," Median " = median(x)
                       , "Q3 " = quantile(x, probs = 0.75)
                       ,"Maximum " = max(x))}
summaryStatsTable <-sapply(c(data[1:1], data[3:4]),stats)

rownames(summaryStatsTable) <- c("Mean", "Stan. Deviation", "Minimum", "Q1", "Median", "Q3", "Maximum")

#number of rows in the data table
rowD <- nrow(data)

ownHomeY = sum(data$ownHome == "ownYes") /rowD
ownHomeN = sum(data$ownHome == "ownNo") /rowD

subY = sum(data$subscribe == "subYes") /rowD
subN = sum(data$subscribe == "subNo") /rowD

mStats <- sum(data$gender == "Male") /rowD
fStats <- sum(data$gender == "Female") /rowD


H0 <- c(fStats, ownHomeN, subN)
H1 <- c(mStats, ownHomeY, subY)


categories <- rbind(H0,H1)
rownames(categories) = c("Female or False", "Male or True")

kables(
  list(
    kable(
      summaryStatsTable, digits = 2, col.names = c("Age","Income", "Kids"), valign = "t", booktabs = "True"
      )
  ,
    kable(
      categories, digits = 3, col.names = c("Gender","OwnsHome", "Subscribes"), valign = "t", booktabs = "True"
      )
  
  ),caption =  "Summary table for all variables") %>%
kable_styling(latex_options = "hold_position")
```
|   As shown by Table 1, ages range between $17$ to $77$ with a median of $39$. The mean _Age_ is $40.74$ which is slightly larger than the median. Additionally, the majority of the ages are in between $33$ and $48.25$. The _Income_ has a very large range, from roughly $11000$ to $139000$. The median, $53000$ and mean, $52000$ are closer to the minimum than the maximum meaning the data is skewed towards the lower values. The 3rd quartile is also very close to the median/mean meaning that there are some potential extreme values. The _Kids_ are also skewed towards lower values. The range goes from $0$ to $7$ but the mean is $1.32$ while the median is $1$. This all gives us an accurate picture of the distribution of the data.
|
|   On the other hand, there is a relatively even proportion between two of the categorical variables, excluding the _subscribers_. For example, while there is roughly $4$ percent difference between males and females, they are both around $50%$. However, the proportion of subscribers is not distributed evenly as more than $85%$ of the sample does not subscribe. While this tells that the majority do not subscribe and that the sample has an even proportion of categories, this does not give us specific categorical values, for example, whether more Females or Males subscribe.


```{r A.1.2 Age histogram, fig.dim = c(3,3), fig.show="hold"}

d1 <- data$age
d2 <- data$income
d3 <- data$kids

hist(d1, breaks=50, main = "Histogram of Age", xlab= "Aggregate age")

hist(d2, breaks=50, main = "Histogram of Income", xlab= "Aggregate income")
```

```{r A.1.2 Kids histogram, fig.dim = c(3,3), fig.show="hold"}
hist(d3, breaks=30, main = "Histogram of Kids", xlab= "Aggregate kids")
```

|   The _Age_ histogram is very centered around the mean while the median is a bit greater than where the mode is. Additionally, there seems to be an observation at $77$ which it could potentially be an outlier. On the other hand,the _Income_ histogram is also centered around its mean and median. There is an observation at roughly $139000$ which is definitely an outlier in this case. Finally, the histogram of _Kids_ is centered around the minimum rather than median or the mean and this could be because of the outlier at $7$. The median in this case is more central than the mean as a result of this outlier. 

```{r A.1.2 Aggregate Boxplot, fig.dim = c(3,3)}

#Aggregate income boxplot
boxplot(d2,
        main = "Aggregate Income boxplot",
        xlab = "Income",
        ylab = "",
        horizontal = TRUE,
        las=1,
        cex.main = 0.9
        )

#Segmented income boxplot
seg_sm<-data[data$Segment=="Suburb mix",]
seg_uh<-data[data$Segment=="Urban hip",]
seg_t<-data[data$Segment=="Travelers",]
seg_mu<-data[data$Segment=="Moving up",]


boxplot(data$income~data$Segment,
        main = "Boxplots for Segmented Income",
        names = c("Suburb Mix", "Urban Hip", "Travelers", "Moving Up"),
        las = 1,
        horizontal = TRUE,
        ylab = "",
        xlab = "Income",
        cex.axis = 0.7,# decrease the font label of the y axis
        cex.main = 0.9
        )

```

|   The _Aggregate Income_ boxplots supports the claims made from the histogram. While there is quite a even distribution centered around the median, there is an extreme value at around $139000$. The quartiles, and the maximum and minimum are also more or less evenly distributed. 
|   As shown by the _Segmented Income_ boxplot, there is a difference between the segments. For example, the median _Moving up Income_ is around $20000$ while for _Travelers_ it is $65000$. Secondly, the _Travelers_ and _Moving Up_ segment have outliers.  
|   Additionally, there is also a much lower minimum and maximum value for Moving Up compared to the rest. The _Moving Up_ and _Urban Hip_ segment is also not skewed compared to the other two. Travelers is slightly skewed towards the left while Suburban mix is skewed to the right. 

|   Do the variables _Income_ and _Age_ appear to be (approximately) normally distributed? Can you think of a suitable distribution for the variable children? 
|
|   The _Income_ and _Age_ variables appear to be approximately normally distributed, even if there appears to be two modes in the _Age_ variable, as mentioned above. With that said,the _Kids_ variable is not approximately normally distributed as it appears rather like a Chi-squared distribution. The distribution is very skewed to the right, with the mode furthest to the left. 


## A.2 - Confidence Intervals

```{r A.2.1, comment= " "}


# references
travelersdata = data[data$Segment == "Travelers",]
urban_hipdata = data[data$Segment == "Urban hip",]
suburb_mixdata = data[data$Segment == "Suburb mix",]
movingUpData <- data[data$Segment == "Moving up",]
maleData <- data[data$gender == "Male",]
femaleData <- data[data$gender == "Female",]

# definitions for income
travelersdata_income <- travelersdata$income
urban_hipdata_income <- urban_hipdata$income
suburb_mixdata_income <- suburb_mixdata$income
movingUpData_income <- movingUpData$income
maleData_income <- maleData$income
femaleData_income <- femaleData$income

income_agg <- data$income
n <- length(income_agg)
m <- mean(income_agg)
s <- sd(income_agg)
z <- qt(0.95, n-1)

ME<-z*s/(n^(0.5))


LCL = m - ME
UCL = m + ME

t <- c(ME,LCL,UCL)

# getting the sample sizes
nx = length(travelersdata_income)
ny = nrow(urban_hipdata)
nz = nrow(suburb_mixdata)
nu = nrow(movingUpData)

# data for travelers
mx <- mean(travelersdata_income)
sx <- sd(travelersdata_income)

# data for urban
my <- mean(urban_hipdata_income)
sy <- sd(urban_hipdata_income)

# data for suburbMix
mz <- mean(suburb_mixdata_income)
sz <- sd(suburb_mixdata_income)

# data for MovingUp
mu <- mean(movingUpData_income)
su <- sd(movingUpData_income)

meT<-z*sx/(nx^(0.5))

lclT <- mx - meT
uclT <- mx + meT

tT <- cbind(meT,lclT,uclT)


meU<-z*sy/(ny^(0.5))

lclU = my - meU
uclU = my + meU

tU <- cbind(meU,lclU,uclU)

meS<-z*sz/(nz^(0.5))

lclS = mz - meS
uclS = mz + meS

tS <- cbind(meS,lclS,uclS)

meM<-z*su/(nu^(0.5))

lclM = mu - meM
uclM = mu + meM




# to get the data from the segmented levels  and male / female look at how we did it in A.4.

```



```{r A.2.1 Aggregate Level, comment= " "}

tTestAgg <- t.test(income_agg, conf.level = 0.90)
tTestTra <- t.test(travelersdata_income, conf.level = 0.90)
tTestUrb <- t.test(urban_hipdata_income, conf.level = 0.90)
tTestSub <- t.test(suburb_mixdata_income, conf.level = 0.90)
tTestMov <- t.test(movingUpData_income, conf.level = 0.90)
tTestMal <- t.test(maleData_income,conf.level = 0.90)
tTestFem <- t.test(femaleData_income, conf.level = 0.90)


aggtTest <- map_df(list(tTestAgg),tidy)
aggtTest[,c("method","alternative")] <- NULL
aggtTest <- cbind("Income",aggtTest)
colnames(aggtTest) <- c("Segments","Estimate","Statistic","P.Value", "DF","LCL","UCL")

MEtTestAgg <- c(aggtTest[,"UCL"]-aggtTest[,"Estimate"])
aggtTest <- cbind(aggtTest,MEtTestAgg)
colnames(aggtTest) <- c("","Estimate","Statistic","P.Value", "DF","LCL","UCL","ME")

kable(aggtTest,
      caption = "T-test table for Aggregate Income, at a 90% C.I for the population mean",
      format="simple", digits = 2,
      )


```

```{r A.2.1 Segment, comment= " "}


finaltTest <- map_df(list(tTestTra,tTestUrb,tTestSub,tTestMov), tidy)



finaltTest[,c("method","alternative")] <- NULL



finaltTest <- cbind(c("Travellers","Urban Hip","Suburb Mix","Moving Up"), finaltTest)
colnames(finaltTest) <- c("Segments","Estimate","Statistic","P.Value", "DF","LCL","UCL")

meTTest <- c(finaltTest[,"UCL"]-finaltTest[,"Estimate"])

finaltTest <- cbind(finaltTest,meTTest)
colnames(finaltTest) <- c("Segments","Estimate","Statistic","P.Value", "DF","LCL","UCL","ME")

kable(finaltTest,
      caption = "T-test table for Segmented Income, at a 90% C.I for the population mean",
      format="simple", digits = 2
      )


#mention that p-value is 0 because it is rounded by the table. It isnt actually 0

```

```{r A.2.1 Gender, comment= " "}

genderTTest <- map_df(list(tTestMal,tTestFem),tidy)

genderTTest[,c("method","alternative")] <- NULL

genderTTest <- cbind(c("Male", "Female"), genderTTest)
colnames(genderTTest) <- c("Gender","Estimate","Statistic","P.Value", "DF","LCL","UCL")

meTTest <- c(genderTTest[,"UCL"]-genderTTest[,"Estimate"])

genderTTest <- cbind(genderTTest,meTTest)
colnames(genderTTest) <- c("Gender","Estimate","Statistic","P.Value", "DF","LCL","UCL","ME")


kable(genderTTest,
      caption = "T-test table for Income by Gender, at a 90% C.I for the population mean",
      format="simple", digits = 2
      )

```

|   We are approximating the sample mean to normal distribution according to the Central Limit Theorem. They are random independent samples, of a size of n $\ge$ 30.
|   The population mean for _Income_ at an aggregate level has a confidence interval with a Lower Confidence Limit (LCL) of 49980.80 and Upper Confidence Limit (UCL) of 53821.89. The population mean lies within this interval with 90% confidence.
|   The population means of _Income_ for these segments lie in their respective confidence intervals that consist of a LCL and an UCL. The segment means lies in their respective intervals with 90% confidence. The _Travelers_ segment has the largest Margin of Error (ME), indicating that the segments mean has the widest spread. Also, since some of the segment confidence intervals overlap, there is no statistically significant difference in _Income_ between the segments. Therefore, the segments cannot be ranked relative to each other in terms of _Income_. In the marketing perspective, this implies that pricing for segments should not only be based on _Income_ but other factors, too.
|   The segments means of _Income_ for males and females lie in their respective confidence intervals that consist of a LCL and UCL The segment means lies in their respective intervals with 90% confidence. According to the confidence intervals and this sample, men have less _Income_ than women. The ME is greater for men, however, so the _Income_ earned varies more for men; therefore, an estimate for the male segment would not necessarily be an accurate representation of real income earned by men. In the marketing perspective, this means that women are willing to pay more for the product, so, to maximize profit potential, Company XYZ should price the product higher for women. The sample data also indicates that the male segment should be sub-segmented further into more economically homogeneous groups such that the pricing strategy towards men is more accurate and effective.


# A.2.2



```{r A.2.2, comment= " ",fig.dim = c(3,3)}

# In this part we are dealing with a true/false variable subsribers (a bool variable either subNo or subYes)
# we want to transform this to numbers so subYes becomes 1 and subNo becomes 0

# at aggregate level we get


n = nrow(data) 
totalP = sum(data$subscribe == "subYes")/n

totalTTestP <- map_df(list(t.test(data$subscribe =="subYes", conf.level = 0.90)),tidy)
totalTTestP[,c("method","alternative")] <- NULL
totalTTestP <- cbind("Subscribers",totalTTestP)

colnames(totalTTestP) <- c("","Estimate","Statistic","P.Value", "DF","LCL","UCL")

MEttestP <- c(totalTTestP[,"UCL"] - totalTTestP[,"Estimate"] )

totalTTestP <- cbind(totalTTestP,MEttestP)
colnames(totalTTestP) <- c("","Estimate","Statistic","P.Value", "DF","LCL","UCL", "ME")

kable(totalTTestP,
      caption ="90% C.I for the population proportion of Subscribers",
        format="simple", digits=2
      )



muN <- nrow(seg_mu)
muP <- sum(seg_mu$subscribe == "subYes")/muN

smN <- nrow(seg_sm)
smP <- sum(seg_sm$subscribe == "subYes") / smN

tN <- nrow(seg_t)
tP <- sum(seg_t$subscribe == "subYes")/tN

uhN <- nrow(seg_uh)
uhP <- sum(seg_uh$subscribe == "subYes") /uhN



proportionC <- rbind(muP,smP,tP,uhP)
rownames(proportionC) <- c("Moving Up", "Suburb Mix", "Travelers", "Urban Hip")


tPropT <- t.test(seg_t$subscribe == "subYes", conf.level = 0.90)
tPropSM <- t.test(seg_sm$subscribe == "subYes", conf.level = 0.90)
tPropMU <- t.test(seg_mu$subscribe == "subYes", conf.level = 0.90)
tPropUH <- t.test(seg_uh$subscribe == "subYes", conf.level = 0.90)

segPTest <- map_df(list(tPropT,tPropSM,tPropMU,tPropUH), tidy)
segPTest[,c("statistic","p.value","parameter","method","alternative")] <- NULL

tTestPME <- c(segPTest[,"conf.high"] - segPTest[,"estimate"])

segPTest <- cbind(c("Travellers","Suburb Mix", "Moving Up", "Urban Hip"),segPTest,tTestPME)

colnames(segPTest) <- c("Segment","Proportion","LCL","UCL", "ME")

kable(segPTest,
      caption = "Population Proportion per Segment with a 90% C.I",
      format ="simple",
      digits = 2
       )

#then look at p 284 in Newbold to figure out the confidence interval 

```

```{r A.2.2 C.I, comment = " ", fig.dim = c(3,3)}


uhTTestP <- t.test(seg_uh$subscribe == "subYes", conf.level = 0.90)

pTable <- map_df(list(uhTTestP),tidy)
pTable[,c("method","alternative")] <- NULL

pTable <- cbind("Urban Hip",pTable)
colnames(pTable) <- c("Segment","Estimate","Statistic","P.Value", "DF","LCL","UCL")

meTTestP <- c(pTable[,"UCL"] - pTable[,"Estimate"] )

pTable <- cbind(pTable,meTTestP)
colnames(pTable) <- c("Segment","Estimate","Statistic","P.Value", "DF","LCL","UCL", "ME")

kable(pTable,
      caption ="90% C.I for the population proportion of Subscribers in the Urban Hip Segment",
        format="simple", digits = 2
      )

```

|   Table 5 shows that the aggregate level estimate shows that only about 13% of the population subscribes to the product, with a ME of $0.03$.  The population proportion of _Subscribers_ for the all segments lies in its confidence interval that consists of a Lower Confidence Limit (LCL) and an Upper Confidence Limit (UCL) with 90% confidence. According to Table 6, _Urban Hip_ has the highest population proportion of _Subscribers_. To maximize profit potential, the Company XYZ should target the Urban Hip segment with marketing strategies aimed at retaining customers in the long-run. To increase the aggregate level of subscription in the population, Company XYZ should execute email campaigns, advertisements, or perhaps a co-creative community to increase the general subscription rate to the product. 


# A.2.3


```{r A.2.3}
age <- data$age
n <- length(age)
s <- sd(age)

# Set the significance level and degrees of freedom
alpha = 0.10
df = n-1  # degrees of freedom = sample size - 1

# Calculate the critical values
lower_crit = qchisq(alpha/2, df)
upper_crit = qchisq(1 - (alpha/2), df)

# Calculate the confidence interval
n <- length(age)  # sample size
s <- sd(age)   # sample standard deviation
LCL = ((df * s^2 / upper_crit)^0.5)
UCL = ((df * s^2 / lower_crit)^0.5)
mePSD <- c(sd(age) - LCL) 


CI <- ((df * s^2 / UCL)^0.5) 
sigma <- ((df * s^2 / LCL)^0.5)

# Print the confidence interval
t<-cbind(s,LCL, UCL, mePSD)
  colnames(t)<-c("Estimate","LCL", "UCL", "ME")
  kable(t,caption = "95% Confidence Interval for Population Age Standard Deviation", 
        format = "simple", digits = 2
    
        )

```

|   Yes, the underlying assumption, as this follows "Case 5: Confidence Intervals for $\sigma^2$ ", is that the data is normally distributed with mean $\mu$ and variance $\sigma^2$. $S^2$ is an estimator for $\sigma^2$, and exactly follows the Chi-Square distribution ($\chi^2$). The estimate is $12.84$ with a ME of ±$0.8$. This concludes a $90$% confidence interval for the population standard deviation from ($12.03;13.77$).

## A.3 - Confidence Intervals and the Sample Size

```{r A.3.1 Part 1}

#3.1 part 1, 2% ME

ME<-0.02
p<-0.5
Z<-qnorm(0.95,0,1)

n1<-((Z^2)*(p*(1-p))/(ME^2))

#maximize p(1-p) -> p=0.5
#n=(z^2*p(1-p))/(ME^2)
#z=qnorm(0.95)
#assume big proportion


#3.1 part 2, 1% ME

ME<-0.01
p<-0.5
Z<-qnorm(0.95,0,1)

n2<-((Z^2)*(p*(1-p))/(ME^2))


#table?

table_3_1<-cbind(n1,n2)
colnames_3_1<-c("2% ME","1% ME")
kable(table_3_1,
      digits=2,
      col.names=colnames_3_1,
      caption="Required sample size to obtain a certain ME",
      format="simple")



```

|   It may not be worth since the sample size would have to increase considerably in order lower the ME only a little bit, in this case from $1691$ to $6764$ (around $4$ times) in order to only decrease the ME from $2$% to $1$%. 

```{r A.3.2, comment = " "}

lNP <- 10000

xP <- totalP * lNP

propTest <- prop.test(xP,
          lNP,
          conf.level = 0.90)

propTestTable <- map_df(list(propTest), tidy)
propTestTable[,c("method","alternative")] <- NULL

propTestTable <- cbind("Aggregate",propTestTable)
colnames(propTestTable) <- c("","Estimate","Statistic","P.Value", "DF","Conf.Low","Conf.High")

propMETTest <- c(propTestTable[,"Conf.High"] - propTestTable[,"Estimate"])
propTestTable <- cbind(propTestTable, propMETTest)
colnames(propTestTable) <- c("","Estimate","Statistic","P.Value", "DF","Conf.Low","Conf.High", "ME")

kable(propTestTable,
      caption = "90% C.I for population proportion of Subscrbers with 10000 samples",
      format = "simple")

```


|   When the sample size increases, the estimation for the population proportion becomes more precise (the ME decreases). As the estimation becomes more precise, it would be possible to increase the confidence level and still get relatively low ME.


## A.4 - Confidence Intervals: Comparison of Population Means


```{r A.4.i}
travelersdata = data[data$Segment == "Travelers",]
urban_hipdata = data[data$Segment == "Urban hip",]


# getting the samplesizes
nx = nrow(travelersdata)
ny = nrow(urban_hipdata)

# data for travelers
mx <- mean(travelersdata$income)
sx <- sd(travelersdata$income)
# data for urban

my <- mean(urban_hipdata$income)
sy <- sd(urban_hipdata$income)

# refer to 8.10 in Newbold p 318

v <- (sx^2/nx + sy^2/ny)^2/ ((sx^2/nx)^2/(nx-1) + (sy^2/ny)^2/(ny-1))

t <- qt(0.95, v)

ME <- t*(sx^2/nx + sy^2/ny)^(0.5)

LCL <- (mx-my) - ME
UCL <- (mx-my) + ME







```

To determine the confidence interval for the difference of population means, we used the Welch's t-test on 80 samples of travelers and 50 samples from the urban hip segment (8.8 Newbold p.317). We assumed that observations in each sample were independently and randomly chosen, and that the populations follow normal distributions. The results showed that the 90% confidence interval for the difference of the mean income $(\bar{x}_{travelers} - \bar{x}_{Urban hip})$ was (42849, 50557), thus indicating a significant difference in the mean income between the two segments with a high degree of confidence.




